A finite-state morphological transducer for Kyrgyz

### 2012-03-13 paper/paper.tex
section worded better, needs additions

\author {Jonathan North Washington\\
Departments of Linguistics and Central Eurasian Studies\\
Indiana University\\
Bloomington, IN 47405-7000 (USA)\\
\texttt{jonwashi@indiana.edu} \and
Mirlan Ipasov\\
Computer Engineering and Mathematics Department\\
International Ataturk Alatoo University\\
Bishkek (Kyrgyzstan)\\
\texttt{mipasov@gmail.com} \and
Francis M. Tyers\\
Dept.\ Lleng.\ i Sist.\ Informàtics \\  
Universitat d'Alacant\\
E-03071 Alacant (Spain)\\
\texttt{ftyers@dlsi.ua.es} 
 }

@@

@@

\subsection{Morphological and orthographic words}
Normal tokenisation often takes white space to be the delimiter between ``words''.  For analysis and generation, however, there are some morphophonological processes that work across the ``whitespace'' boundary.

>>>>

\author {Jonathan North Washington\\
Departments of Linguistics and Central Eurasian Studies\\
Indiana University\\
Bloomington, IN 47405 (USA)\\
\texttt{jonwashi@indiana.edu} \and
Mirlan Ipasov\\
Computer Engineering and Mathematics Department\\
International Ataturk Alatoo University\\
Bishkek (Kyrgyzstan)\\
\texttt{mipasov@gmail.com} \and
Francis M. Tyers\\
Dept.\ Lleng.\ i Sist.\ Informàtics \\  
Universitat d'Alacant\\
E-03071 Alacant (Spain)\\
\texttt{ftyers@dlsi.ua.es} 
 }

@@

\begin{abstract}
Less than 200 words

\end{abstract}

@@

\subsection{Morphological and orthographic words}
Normal tokenisation often takes white space to be the delimiter between ``words''.  For analysis and generation, however, there are two exceptions to this.  Some morphophonological processes work across the ``whitespace'' boundary, and some clitics which are written next to the previous word without a whitespace can be considered syntactically separate units (words) and follow standard morphophonologial processes.

Kyrgyz, like other turkic languages (Tatar, Chuvash, Kazakh), has morphophonological processes that work across the ``whitespace'' boundary, but unlike many other Turkic languages, they are not represented in the orthography.

Some clitics in Kyrgyz which can be considered separate words but are written together with the previous word include the question word `бы'; the focus particle `чы'; copula suffixes `мин', etc.; and the progressive auxiliary `жат' with certain verbs.
%FIXME: why copular suffixes?

One result of making transducer targetted at machine translation is that we want to be able to both analyse, and generate the same forms. .... %FIXME
